---
title: "Notions of Racism in STEM"
subtitle: "A Science Mapping Analysis"
author: 
  - "Nathan Alexander, PhD (Howard University)"
  - "Qyana Stewart, PhD student (Howard University)"
  - "Basil Ghali, PhD student (Howard University)"
output:
  html_document:
    theme: flatly
  pdf_document: default
editor_options:
  markdown:
    wrap: sentence
---

# Overview

We conduct a systematic review of journal articles which examine racism in relation to the sciences, technology, engineering, and mathematics (STEM). The term 'notion' is used to describe context-dependent instances of racism that point to broader meanings, that exist across contexts, which go on to formulate the evolving understandings of racism, noticeable in different ways that scholars have defined racism across disciplines and different settings.

## Conceptual Framework

This study is motivated, in part, by the work of Shiao & Woody (2021) who discuss the various meanings of racism in sociological journal content. For the purpose of this study, our team sought to make sense of the nature of this intellectual structure in STEM, and the various concepts and themes that are situated within this structure.

| **Bonilla-Silva** | **Shiao and Woody**                 | **Notions** |
|-----|-------------|---------|
| _Pyschological_   | Individual attitudes                | Personal    |
| _Cultural_        | Cultural schema                     | Cultural    |
| _Structural_      | Structural: Pre-existing conditions | Historical  |
|                   | Structural: Create or maintain      | Futural     |
|                   | Structural: Spatial                 | Local       |
|                   |                                     | Global      |

### Replication

#### Bibliographic collection

*Data source*: Clariate Analytics Web of Science

*Data format*: Plain text

*Time span*: 2003 - 2023

*Document Type*: Articles; Articles, early access

*Query date*: July 1, 2024


### Research question

*What is the state of the intellectual structure and emerging notions in research on racism in STEM?*

## Methodological Framework

We integrate systematic review techniques with computational methods to generate the methods for the study. The systematic review process builds on qualitative meta-synthesis techniques, as well as content and thematic analysis. The computational methods follow the general scientific mapping workflow for working with large sets of citations data.

### Data

Four databases were used initially to identify the possible collections of sources, and to compare the various results obtained across databases.

**ERIC**.

**Scopus**.

**Google Scholar**.

**Web of Science (WoS)**. The WoS Core Collection was used to obtain bibliographic information for the study sample. The Science Citation Index Expanded (SCI-EXPANDED), Social Science Citation Index (SSCI), Arts & Humanities Citation Index (ACHI), and Emerging Sources Citation Index (ESCI) were selected. The Conference Proceedings Citation Index - Science (CPCI-S), Conference Proceedings Citation Index - Social Science (CPCI-SSH), Book Citaton Index - Science (BKCI-S), and Book Citation Index - Social Science & Humanities (BKCI-SSH) were not selected.

Given the formal inclusion of psychology as one of the STEM disciplines by research foundations, such as the National Science Foundation (NSF) in the United States, social science bibliographic entries were included.

[Query link]{https://www.webofscience.com/wos/woscc/summary/6d0d4c0d-2e83-4162-bd8f-077ad0a4cc34-f7d6e8a1/relevance/1}

A replication of the Shian & Woody results was obtained to ensure the appropriate steps for the conceptual replication were derived. The initial search for articles from the Web of Science Core Collection included the query (((TI=(racism or racist)) OR AB=(racism or racist)) OR AK=(racism or racist)) AND DOP=(1995-01-01/2015-12-31)) AND LA=(English) returned 8,950 results. When the article requirement and individual collection indices were added (SSCI, A&HCI, ESCI, and SCI-EXPANDED), a total of 6,378 articles remained. When the results were subset to those indexed as Sociology, a total of 1,033 articles remained.

(TI=(racism or racist) OR AB=(racism or racist) OR AK=(racism or racist)) AND DOP=(1995-01-01/2015-12-31) AND LA=(English)

#### Year span: 1995-2015

When education journal articles were selected, a total of 853 results remained.
https://www.webofscience.com/wos/woscc/summary/bfe60ccd-86c6-40d0-9ba5-16b667ff199e-f83fbe53/relevance/1 

When science, technology, engineering, and mathematical subjects were selected a total of 1,462 results remained.
https://www.webofscience.com/wos/woscc/summary/eff6a28e-2893-454e-a79f-13e94a7d6c0e-f83fba7d/relevance/1

Forty results were returned when the sample is subset to education journal articles including one of the terms science, technology, engineering, mathematics, or STEM in their title or abstract.
https://www.webofscience.com/wos/woscc/summary/d5c643f1-25c7-46c1-93ef-6f458652d262-f83fe3c7/relevance/1


#### Year span: 2016-2023


#### Analytic sample
A final analytic sample of 254 articles was selected.
These articles included the words racism or racist in their abstracts and at least one of the terms, science, technology, engineering, mathematics, or STEM in their abstracts); the time period was expanded to include article for the 20 year period from 2003 to 2023.
https://www.webofscience.com/wos/woscc/summary/331dedce-0c5c-40cd-b63e-734077ea3736-f840d6cf/relevance/1


**Results**.

Final data analysis was run on June 28, 2024. Prior to this period, multiple programs were designed and checked to ensure accuracy across different samples and generated psuedo data. All programs were confirmed to return the same results across studies.

3,318,115 results from Science Citation Index Expanded (SCI-EXPANDED), Social Sciences Citation Index (SSCI), Arts & Humanities Citation Index (A&HCI), Emerging Sources Citation Index (ESCI): for ABSTRACT AND racism or science or technology or engineering or mathematics or stem
https://www.webofscience.com/wos/woscc/summary/b7144fe8-2337-4d15-b2db-0fc78da08022-f78ebcfe/relevance/1


## Analytic Framework

```{r, echo=F, eval=T, fig.align='center'}
library(DiagrammeR)

grViz(" digraph model {

    graph [layout = dot, overlap = true]
    
    node [shape=box]
    Scoping -> Comparison -> Reduction;
 
    node [shape=box]
    data [label=<Bibliographic Entries>];
    
    node [shape=box]
    format [label=<Document Matrix>];
    
# `edge` statements
 
    edge [minlen = 2]
    data -> format
    format -> {Concepts Themes} 
    {Concepts Themes} -> Notions
    Meanings -> Notions
 
 }") 
```

# Set up

Load packages and import for data analysis.

### Load packages

```{r, echo=T, eval=T, warning=F}
library(dplyr)
library(readtext)
library(tidyverse)
library(here)
library(gt)
here::i_am("analysis.Rmd")
```

### Set up `Bibliometrix`

```{r, echo=T, eval=T, warning=F}
# install.packages("bibliometrix")
library(bibliometrix)

# test run biblioshiny
# bibliometrix::biblioshiny()
```

### Load and convert data

```{r, echo=T, eval=T, warning=F}
# load wos database results
edrecords <- readtext(here("final-sample-ed.txt"))
# https://www.webofscience.com/wos/woscc/summary/331dedce-0c5c-40cd-b63e-734077ea3736-f840d6cf/relevance/1

stemrecords <- readtext(here("final-sample-stem.txt"))
# https://www.webofscience.com/wos/woscc/summary/507ef07d-4440-43c5-a139-c3148bf75a7f-f8416bcd/relevance/1

# convert  file to a bibliometric df
M1 <- convert2df(file=edrecords, dbsource= "wos", format= "plaintext")
M2 <- convert2df(file=stemrecords, dbsource= "wos", format= "plaintext")
```

#### Notes

  - The file contains 261 documents from lsi. 
  
  - The file is missing the `Cited References` (CR) as noted above; the CR were omitted as a part of the data generation process. 
  
  - The status of the DE (Keyword) category is labeled as `Poor` given that 62 missing counts (prop=0.2375) are listed; however, the ID (Keywords plus) category is listed as `Acceptable`, with only 30 missing counts (prop=0.1149). 
  
  - All other indicators are listed as `Good` (at least 1 missing count) or `Excellent` (0 missing counts).

# Descriptive Analysis

Although bibliometrics is primarily used to understand productivity and the impact of specific contributions to a field, the field has expanded to inform useful insights in systematic analysis.

### Main information about the included studies

```{r}
results1 <- biblioAnalysis(M1)
summary(results1, k=20, pause=F, width=130)
plot(x=results1, k=10, pause=F)

results2 <- biblioAnalysis(M2)
summary(results2, k=20, pause=F, width=130)
plot(x=results2, k=10, pause=F)
```


## Most Cited References

Results across the two samples indicate that the two disciplinary structures pull from very different sources of literature. The source of literature in the education sciences is highly concentrated whereas the literature in the STEM fields is more distributed. 

Kimberlee Crenshaw's 1993 Stanford Law Review article is the highest citation that shows up on both lists.

Researchers publishing in STEM outlets tend to not cite studies from education research. Studiens in education reserach, similarly, tend to pull from other studies in education.

```{r}
CR1 <- citations(M1, field = "article", sep = ";")
cbind(CR1$Cited[1:20])

CR2 <- citations(M2, field = "article", sep = ";")
cbind(CR2$Cited[1:20])
```


# Intellectual Structure

## Co-citation analysis

```{r}
NetMatrix1 <- biblioNetwork(M1, analysis = "co-citation", network = "references", sep = ";")
net1=networkPlot(NetMatrix1, n = 50, Title = "Co-Citation Network", type = "fruchterman", size.cex=TRUE, size=14, remove.multiple=T, labelsize=1,edgesize = 7, edges.min=4)
```

```{r}
NetMatrix2 <- biblioNetwork(M2, analysis = "co-citation", network = "references", sep = ";")
net2=networkPlot(NetMatrix2, n = 50, Title = "Co-Citation Network", type = "fruchterman", size.cex=TRUE, size=14, remove.multiple=T, labelsize=1,edgesize = 7, edges.min=4)
```

### Descriptive analysis of co-citation network characteristics

```{r}
netstat1 <- networkStat(NetMatrix1)
summary(netstat1,k=20)

netstat2 <- networkStat(NetMatrix2)
summary(netstat2,k=20)
```


## Journal (source) co-citation analysis

```{r}
M1=metaTagExtraction(M1,"CR_SO",sep=";")
NetMatrix1 <- biblioNetwork(M1, analysis = "co-citation", network = "sources", sep = ";")
net1=networkPlot(NetMatrix1, n = 30, Title = "Co-Citation Network", type = "auto", size.cex=TRUE, size=15, remove.multiple=FALSE, labelsize=1,edgesize = 10, edges.min=5)

M2=metaTagExtraction(M2,"CR_SO",sep=";")
NetMatrix2 <- biblioNetwork(M2, analysis = "co-citation", network = "sources", sep = ";")
net2=networkPlot(NetMatrix2, n = 30, Title = "Co-Citation Network", type = "auto", size.cex=TRUE, size=15, remove.multiple=FALSE, labelsize=1,edgesize = 10, edges.min=5)

```

# Historiograph

```{r}
histResults1 <- histNetwork(M1, sep = ";")
options(width = 130)
net1 <- histPlot(histResults1, n=20, size = 5, labelsize = 4)

histResults2 <- histNetwork(M2, sep = ";")
options(width = 130)
net2 <- histPlot(histResults2, n=20, size = 5, labelsize = 4)
```

# Conceptual structre

## Co-word analysis

### Keyword occurences

```{r}

NetMatrix1 <- biblioNetwork(M1, 
                            analysis = "co-occurrences", 
                            network = "keywords", 
                            sep = ";")
net1=networkPlot(NetMatrix1, 
                 normalize="association", 
                 n = 50, 
                 Title = "Keyword Co-occurrences", 
                 type = "fruchterman", 
                 size.cex=TRUE, 
                 size=10, 
                 remove.multiple=F, 
                 edgesize = 10, 
                 labelsize=3,
                 label.cex=TRUE,
                 label.n=15,
                 edges.min=2)


NetMatrix2 <- biblioNetwork(M2, 
                            analysis = "co-occurrences", 
                            network = "keywords", 
                            sep = ";")
net2=networkPlot(NetMatrix2, 
                 normalize="association", 
                 n = 50, 
                 Title = "Keyword Co-occurrences", 
                 type = "fruchterman", 
                 size.cex=TRUE, size=10, 
                 remove.multiple=F, 
                 edgesize = 10, 
                 labelsize=3,
                 label.cex=TRUE,
                 label.n=15,
                 edges.min=2)

```

Descriptive analysis of the keyword co-occurence network characteristics.

```{r}
netstat1 <- networkStat(NetMatrix1)
summary(netstat1,k=10)

netstat2 <- networkStat(NetMatrix2)
summary(netstat2,k=10)

```

### Correspondence analysis
```{r}
suppressWarnings(
CS1 <- conceptualStructure(M1, method="MCA", field="ID", minDegree=15, clust=5, stemming=FALSE, labelsize=8,documents=20)
)

suppressWarnings(
CS2 <- conceptualStructure(M2, method="MCA", field="ID", minDegree=15, clust=5, stemming=FALSE, labelsize=8,documents=20)
)
```

# Thematic map

```{r}
Map1=thematicMap(M1, field = "ID", n = 250, minfreq = 4,
  stemming = FALSE, size = 0.7, n.labels=5, repel = TRUE)
plot(Map1$map)


Map2=thematicMap(M2, field = "ID", n = 250, minfreq = 4,
  stemming = FALSE, size = 0.7, n.labels=5, repel = TRUE)
plot(Map2$map)

```

Cluster descriptions

```{r}
library(dplyr)

Clusters1=Map1$words[order(Map1$words$Cluster,-Map1$words$Occurrences),]
Clusters2=Map2$words[order(Map2$words$Cluster,-Map2$words$Occurrences),]

CL1 <- Clusters1 %>% group_by(.data$Cluster_Label) %>% top_n(5, .data$Occurrences)
CL1

CL2 <- Clusters2 %>% group_by(.data$Cluster_Label) %>% top_n(5, .data$Occurrences)
CL2
```
---

## **References**

Aria, M. & Cuccurullo, C. (2017) bibliometrix: An R-tool for comprehensive science mapping analysis, *Journal of Informetrics, 11*(4), pp 959-975, Elsevier.

Aria, M., Cuccurullo, C., D’Aniello, L., Misuraca, M., & Spano, M. (2022). Thematic Analysis as a New Culturomic Tool: The Social Media Coverage on COVID-19 Pandemic in Italy. *Sustainability, 14*(6), 3643. https://doi.org/10.3390/su14063643.

Aria M., Misuraca M., Spano M. (2020) Mapping the evolution of social research and data science on 30 years of Social Indicators Research, *Social Indicators Research*. https://doi.org/10.1007/s11205-020-02281-3


Cobo, M. J., Lopez-Herrera, A. G., Herrera-Viedma, E., & Herrera, F. (2011). An approach for detecting, quantifying, and visualizing the evolution of a research field: A practical application to the fuzzy sets theory field. *Journal of Informetrics, 5*(1), 146-166.

Shiao & Woody (2019). The meaning of racism. *Sociological Perspectives, 12*(X), pp 100-101, Oxford.