---
title: "Overview: Keyword `racism`"
author: "Nathan Alexander"
date: "2024-02-25"
output:
  html_document:
    toc: true
    self_contained: yes
    mode: selfcontained
  pdf_document:
    keep_tex: true
---

# Google trend analysis of racism

We generate a basic analysis of Google Trends data on the term `racism`.

```{r load-packages, include=FALSE}
library(dplyr)
library(magrittr)
library(knitr)

install.packages("gtrendsR", repos = "http://cran.us.r-project.org")
# load libraries
library(dplyr)
library(magrittr)
library(gtrendsR)
library(knitr)
library(ggplot2)
library(patchwork)
```

```{r, include=FALSE}
# gather country codes and names
data(countries)
countries %>% 
  filter(country_code == "US")
```

## 10-year Google Trend 
Searches for the term `racism` in the United States

```{r}
# racism (10-year trend)
racism10year<-gtrends(c("racism"), time= "2013-01-01 2023-12-31", geo = "US")
plot(racism10year)
```

Hypothesis was a spike in US between 2019 and 2022.

## 2-year Google Trend for 2020 and 2021
Searches for the term `racism` in the United States

```{r}
# racism (1-year trend for 2021)
racism1year<-gtrends(c("racism"), time= "2019-12-31 2021-12-31", geo = "US")
plot(racism1year)
```

## Search for `racism` and `covid`
Adding 10-year historical context (Covid) to searches for racism

```{r}
racism_covid<-gtrends(c("covid", "racism"), time= "2013-01-01 2023-12-31", geo = "US")
plot(racism_covid)
```

## Search for `racism` and `stem`
Adding 10-year ancillary context (subject=stem) to searches for racism

```{r}
racism_stem<-gtrends(c("stem", "racism"), time= "2013-01-01 2023-12-31", geo = "US")
plot(racism_stem)
```

## Search for `racism` and `education`
Adding 10-year ancillary context (subject=education) to searches for racism

```{r}
racism_education<-gtrends(c("education", "racism"), time= "2013-01-01 2023-12-31", geo = "US")
plot(racism_education)
```

## Search for `racism` and `george floyd`
Adding historical context (person=george floyd) to searches for racism

```{r}
racism_george_floyd<-gtrends(c("george floyd", "racism"), time= "2019-12-31 2021-12-31", geo = "US")
plot(racism_george_floyd)
```

```{r}
racism_george_floyd<-gtrends(c("george floyd", "racism"), time= "2019-12-31 2021-12-31", geo = "US")

racism_george_floyd_2 <- racism_george_floyd$interest_over_time %>% 
  dplyr::mutate(hits=ifelse(hits=="<1",0.5, as.numeric(hits)),
         date=as.Date(date))

ggplot(racism_george_floyd_2, aes(x=date, y=hits, color=keyword)) +
  geom_line() +
  geom_vline(xintercept=as.numeric(as.Date("2020-05-25")))
```

## Add date of George Floyd murder by police
Layering date of George Floyd's murder (may 25, 2020) to plot.

```{r}
racism1year<-gtrends(c("racism", "george floyd", "covid"), time= "2019-12-31 2021-12-31", geo = "US")

racism1year_trend <- racism1year$interest_over_time %>% 
  dplyr::mutate(hits=ifelse(hits=="<1",0.5, as.numeric(hits)),
         date=as.Date(date))

ggplot(racism1year_trend, aes(x=date, y=hits, color=keyword)) +
  geom_line() +
  geom_vline(xintercept=as.numeric(as.Date("2020-05-25")))
```

The above information will inform our keyword inquiries for a bibliometric analysis.
